---
title: "Analisa Data Covid-19 di Provinsi Jawa Barat"
author: "Roby"
date: "8/3/2020"
output:
  html_document: default
  pdf_document: default
---

library(httr)
library(hrbrthemes)
library(ggplot2)
library(dplyr)
library(lubridate)
library(tidyr)

Optional: use this if you got warning messages `font family not found in your database`
library(extrafont)
-->> also this: font_import()
quit and re-open your session

## Access the API from covid.go.id for the province of West Java

```{r}
resp <- GET("https://data.covid19.go.id/public/api/prov_detail_JAWA_BARAT.json")
```

## Extract the response content

```{r}
cov_jabar_raw <- content(resp, as = "parsed", simplifyVector = TRUE)
```

## Check component name

```{r}
names(cov_id_raw)
```

```{r}
head(cov_jabar_raw)
```

### What is the total number of COVID-19 cases in West Java?

```{r}
cov_jabar_raw$kasus_total
```

### What percentage of deaths due to COVID-19 in West Java?

```{r}
cov_jabar_raw$meninggal_persen
```

### What is the percentage recovery rate of COVID-19 in West Java?

```{r}
cov_jabar_raw$sembuh_persen
```

## Save the history of the covid-19 development of West Java province into a new variable

```{r}
cov_jabar <- cov_jabar_raw$list_perkembangan
```

```{r}
str(cov_jabar)
```
```{r}
head(cov_jabar)
```

## Data Manipulation and save into `new_cov_jabar`

1. Deleting columns "DIRAWAT_OR_ISOLASI" and "ACCUMULASI_DIRAWAT_OR_ISOLASI"
2. Delete all columns that contain cumulative values
3. Changing the name of the column `KASUS` to `kasus_baru`
4. Change the format of writing the following column to lowercase
     - column `MENINGGAL`
     - column `SEMBUH`
5. Correcting data in the date column

```{r}
new_cov_jabar <- 
  cov_jabar %>%
  select(-contains("DIRAWAT_OR_ISOLASI")) %>%
  select(-starts_with("AKUMULASI")) %>% 
  rename(
    kasus_baru = KASUS,
    meninggal = MENINGGAL,
    sembuh = SEMBUH
  ) %>% 
  mutate(
    tanggal = as.Date.POSIXct(tanggal/1000, origin = "1970-01-01"),
    tanggal = as.Date(tanggal)
  )

str(new_cov_jabar)
```
## Data Visualization

```{r}
ggplot(new_cov_jabar, aes(tanggal, kasus_baru)) +
  geom_col(fill = "salmon") +
  labs(
    x = NULL,
    y = "Jumlah kasus",
    title = "Kasus Harian Positif COVID-19 di Jawa Barat",
    subtitle = "Terjadi pelonjakan kasus di awal bulan Juli akibat klaster Secapa AD Bandung",
    caption = "Sumber data: covid.go.id"
  ) +
  theme_ipsum(
    base_size = 13,
    plot_title_size = 21,
    grid = "Y",
    ticks = TRUE
  ) +
  theme(plot.title.position = "plot")
```

```{r}
ggplot(new_cov_jabar, aes(tanggal, sembuh)) +
  geom_col(fill = "olivedrab2") +
  labs(
    x = NULL,
    y = "Jumlah kasus",
    title = "Kasus Harian Sembuh COVID-19 di Jawa Barat",
    subtitle = "Terjadi pelonjakan kasus di awal bulan Juli akibat klaster Secapa AD Bandung",
    caption = "Sumber data: covid.19.go.id"
    ) +
  theme_ipsum(
    base_size = 13,
    plot_title_size = 21,
    grid = "Y",
    ticks = TRUE
    ) +
  theme(plot.title.position = "plot")
```

```{r}
ggplot(new_cov_jabar, aes(tanggal, meninggal)) +
  geom_col(fill = "darkslategray4") +
  labs(
    x = NULL,
    y = "Jumlah kasus",
    title = "Kasus Harian Meninggal Akibat COVID-19 di Jawa Barat",
    caption = "Sumber data: covid.19.go.id"
  ) +
  theme_ipsum(
    base_size = 13, 
    plot_title_size = 21,
    grid = "Y",
    ticks = TRUE
  ) +
  theme(plot.title.position = "plot")
```

## How is the case developing over a weekly period?


```{r}
cov_jabar_pekanan <- new_cov_jabar %>% 
  count(
    tahun = year(tanggal),
    pekan_ke = week(tanggal),
    wt = kasus_baru,
    name = "jumlah"
  )

glimpse(cov_jabar_pekanan)
```

### Is this week better than last week?

1. Create a new column that contains the number of new cases in the previous week. This column is named `jumlah_pekanlalu`
2. Replace the NA value in the `jumlah_pekanlalu` with the value `0`
3. Make a comparison between the column `jumlah` with the column `jumlah_pekanlalu`. The results of this comparison are stored in a new column with the name `lebih_baik`, the value is `TRUE`if the number of new cases this week is lower than the number of cases last week

```{r}
cov_jabar_pekanan <-
  cov_jabar_pekanan %>% 
  mutate(
    jumlah_pekanlalu = dplyr::lag(jumlah, 1),
    jumlah_pekanlalu = ifelse(is.na(jumlah_pekanlalu), 0, jumlah_pekanlalu),
    lebih_baik = jumlah < jumlah_pekanlalu
  )

glimpse(cov_jabar_pekanan)
```

## Make a Bar Chart

```{r}
ggplot(cov_jabar_pekanan, aes(pekan_ke, jumlah, fill = lebih_baik)) +
  geom_col(show.legend = FALSE) +
  scale_x_continuous(breaks = 9:31, expand = c(0, 0)) +
  scale_fill_manual(values = c("TRUE" = "seagreen3", "FALSE" = "salmon")) +
  labs(
    x = NULL,
    y = "Jumlah kasus",
    title = "Kasus Pekanan Positif COVID-19 di Jawa Barat",
    subtitle = "Kolom hijau menunjukan penambahan kasus baru lebih sedikit dibandingkan satu pekan sebelumnya",
    caption = "Sumber data: covid.19.go.id"
    ) +
  theme_ipsum(
    base_size = 13,
    plot_title_size = 21,
    grid = "Y",
    ticks = TRUE
    ) +
  theme(plot.title.position = "plot")
```

## Until now, how many cases are still active? Active in the sense of being treated or isolated.

```{r}
cov_jabar_akumulasi <- 
  new_cov_jabar %>% 
  transmute(
    tanggal,
    akumulasi_aktif = cumsum(kasus_baru) - cumsum(sembuh) - cumsum(meninggal),
    akumulasi_sembuh = cumsum(sembuh),
    akumulasi_meninggal = cumsum(meninggal)
  )

tail(cov_jabar_akumulasi)
```

## Make a Line Chart

```{r}
ggplot(data = cov_jabar_akumulasi, aes(x = tanggal, y = akumulasi_aktif)) +
  geom_line()
```

## Pivoting the `cov_jabar_akumulasi` data, then creating a `geom_line()` layer using aesthetic colors


```{r}
dim(cov_jabar_akumulasi)

cov_jabar_akumulasi_pivot <- 
  cov_jabar_akumulasi %>% 
  gather(
    key = "kategori",
    value = "jumlah",
    -tanggal
  ) %>% 
  mutate(
    kategori = sub(pattern = "akumulasi_", replacement = "", kategori)
  )

dim(cov_jabar_akumulasi_pivot)

glimpse(cov_jabar_akumulasi_pivot)
```

### Since `tidyr` version 1.0.0, you are advised to use the `pivot_longer()` function instead of `gather()` and `pivot_wider()` instead of `spread()`. `pivot_longer()` and `pivot_wider()` have more complete features than `gather()` and `spread()`. The transformation process of `cov_jabar_akumulasi` into `cov_jabar_akumulasi_pivot` can be done using `pivot_longer()` as follows:

```{r}
cov_jabar_akumulasi_pivot2 <- 
  cov_jabar_akumulasi %>% 
  pivot_longer(
    cols = -tanggal,
    names_to = "kategori",
    names_prefix = "akumulasi",
    values_to = "jumlah"
  )

dim(cov_jabar_akumulasi_pivot)

glimpse(cov_jabar_akumulasi_pivot)
```

## Make a comparison chart between the accumulation of active cases `kasus_aktif`, cured cases `kasus_sembuh` and death cases `kasus_meninggal`

```{r}
ggplot(cov_jabar_akumulasi_pivot, aes(tanggal, jumlah, colour = (kategori))) +
  geom_line(size = 0.9) +
  scale_y_continuous(sec.axis = dup_axis(name = NULL)) +
  scale_colour_manual(
    values = c(
      "aktif" = "salmon",
      "meninggal" = "darkslategray4",
      "sembuh" = "olivedrab2"
      ),
    labels = c("Aktif", "Meninggal", "Sembuh")
    ) +
  labs(
    x = NULL,
    y = "Jumlah kasus akumulasi",
    colour = NULL,
    title = "Dinamika Kasus COVID-19 di Jawa Barat",
    caption = "Sumber data: covid.19.go.id"
    ) +
  theme_ipsum(
    base_size = 13,
    plot_title_size = 21,
    grid = "Y",
    ticks = TRUE
    ) +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "top"
    )
```




